The Deviation of China Map as a Regression Problem

### 2012-09-09 paper.tex
upate

\section{Notation}
We need to find out two functions, $f_x(x,y)$ and $f_y(x,y)$, to describe the
deviation. $x$ and $y$ are the true longitude and latitude respectively.
$f_x(x,y)$ and $f_y(x,y)$ are the deviation at point $(x,y)$. In other words,
if $(x,y)$ is the deviated coordinate, then $(f_x(x,y)+x,f_y(x,y)+y)$ is the
deviated coordinate.
Since $|f_x(x+d,y)-f_x(x,y)|$ is much smaller than $d$, and same for $y$,
we can also
use the two functions to calculate the true coordinate from a deviated
coordinate, i.e. if $(x,y)$ is the deviated coordinate, then
$(x-f_x(x,y),y-f_y(x,y))$ is the true coordinate.

In the following analysis, we will use octave.

In my previous post, ... we first solve the functions in one dimention by setting one parameter to be constant.

@@

\noindent
\verb|> data=load("-ascii", "iny-f-110.txt");| \\
\verb|> y=data(:,2);| \\
\verb|> dx=data(:,3);| \\
\verb|> plot(y,dx)       | \hfill Figure~\ref{fig:fxy-degree}(left)

@@

So the function is $f_y(x,39.36)=-160.45195+2.42764x+13.33287sin(2\pi
x)-0.27150cos(2\pi x)+13.30099cos(6\pi x)-0.82075cos(6\pi x)$.
We can combine sin and cos of the same frequency.

\begin{verbatim}
> atan(0.82075/13.30099)
ans =  0.061628
> atan(0.27150/13.33287)
ans =  0.020360
> mx=[ones(size(x)),x,sin(2.*pi.*x-0.020360),sin(6.*pi.*x-0.061628)];
> std(mx*(pinv(mx)*dy)-dy)
ans =  0.38114
> pinv(mx)*dy
ans =
  -160.4519
     2.4276
    13.3356
    13.3263
\end{verbatim}
So the function is $f_y(x,39.36)=-160.45195+2.42764x+13.3356sin(2\pi
x-0.061628)+13.3263cos(6\pi x-0.020360)$.

After trying out other constant latitudes, we find out that the first two
coefficients to be changing but the last 4 coefficients to be fixed.
Thus, we know
$f_y(x,y)=g_1(y)+g_2(y)x+13.3356sin(2\pi x-0.061628)+13.3263cos(6\pi
x-0.020360)$, where $g_1(y)$ and $g_2(y)$ are not yet known.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.48\textwidth]{fxx-lc2.png}
\end{center}
\caption{9-order polynomial (green) and 9-order polynomial + sinusoidal (blue)}
\label{fig:fxx-lc2}
\end{figure}

Although the 9-order polynomial fits the data well (error is 0.79 meters),
the function is ugly.
After lots of tries\footnote{
I have spent most time here, but there is nothing to say about it because it's
simply trying different functions.
There are still a little bit of technique though.
For example, the two green peaks in Figure~\ref{fig:fxx-lc2} have distance
about 24 in x-axis, thus there might be a sinusoidal component of period 24.
}, I figured out that the low frequency
component actually consists of a quadratic curve and two sinusoidal wave of
period 60 and 24.
Using this model, the regression error is 0.39 meters and the function is
simple and beautiful.
\begin{verbatim}
> mx=[ones(size(x)),x,x.*x,sin(pi.*x./30),cos(pi.*x./30),\
  sin(pi.*x./12),cos(pi.*x./12),sin(pi.*x./3),cos(pi.*x./3),\
  sin(pi.*x),cos((pi.*x)),sin(2.*pi.*x),cos(2.*pi.*x),sin(6.*pi.*x),\
  cos((6.*pi.*x))]; std(mx*(pinv(mx)*dx)-dx)
ans =  0.38668
> pinv(mx)*dx
ans =
   1.2052e+03
  -1.8404e+01
   9.4022e-02
   8.3824e-01
   2.0044e+02
  -7.0435e+01
  -7.0449e+01
  -2.6699e+01
   9.1218e-02
  -1.3342e+01
   1.3044e-01
   1.3342e+01
  -2.7504e-01
   1.3306e+01
  -8.2298e-01
\end{verbatim}

@@

Finally, $f_x(x,39.36) = 1205.1 -18.402x +0.094012x^2 +200.45sin(\pi
x/30+1.5666) -99.62sin(\pi x/12+0.78550) -26.699sin(\pi x/3-0.0034165)
-13.343sin(\pi x-0.0097763) +13.344sin(2\pi x-0.020612) +13.332sin(6\pi
x-0.061772)$. To see how the components fit the data, we can graduately add
the higher frequency components. Figure 5.5 shows this this idea. Figure 5.6 is
a zoomed in region in Figure 5.5. The first function (in red) is the quadratic
function $1205.1 -18.402x +0.094012x^2$. The second function (in green) adds
one sinusoidal component to it: $1205.1 -18.402x +0.094012x^2 +200.45sin(\pi
x/30+1.5666)$. The third function (in blue) adds one more sinusoidal component
to it, etc.
\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.48\textwidth]{fxx-lc3.png}
\includegraphics[width=0.48\textwidth]{fxx-lc4.png}
\end{center}
\caption{Decomposition of Components.
The first function (red) has only the lowest frequency component.
The second function (green) adds the second lowest frequency component to the
first function.
The third (blue) adds another component, etc.
Right side is a zoomed-in region in right side.}
\label{fig:fxx-lc3}
\end{figure}

After trying out other constant latitude, we found that only the first two
coefficients are affected by latitude. The rest of the coefficients do not
change. Thus, we can write $f_x(x,y)=g_1(y) -g_2(y)x +0.094012x^2
+200.45sin(\pi x/30+1.5666) -99.62sin(\pi x/12+0.78550) -26.699sin(\pi
x/3-0.0034165) -13.343sin(\pi x-0.0097763) +13.344sin(2\pi x-0.020612)
+13.332sin(6\pi x-0.061772)$, where $g_1(y)$ and $g_2(y)$ are not yet known.

@@

The process of finding $f_y(const, y)$ is similar to $f_x(x, const)$, so I skip
it. The function is similar to $f_x(x, const)$ except that the two high
frequency sinusoidal components do not appear in $f_y(const, y)$.

\section{The final 2D case: $f_x(x,y)$ and $f_y(x,y)$}
With out too much effort, we have found out the two functions:
$f_x(x,y) = 1598.1-23.605x -8.4345y +0.09933xy +0.10069x^2
+199.1cos(\pi(x+0.017)/30) -99.772sin(\pi(x+0.0039)/12+\pi/4)
-26.687sin(\pi(x-0.0039)/3) -13.345sin(\pi(x-0.0039))
+13.338sin(2\pi(x-0.0039)) +13.331sin(6\pi(x-0.0039))$ $f_y(x,y) = -36.759
+1.5218x +10.002y -0.10032xy -0.03661y^2 -217.44cos(\pi(y+8.143)/30)
+106.41sin(\pi(y+1.1)/12) -26.697sin(\pi(y+1)/3) +13.335sin(\pi y)
-13.329sin(2\pi(x-0.0039)) -13.319sin(6\pi(x-0.0039))$
The standard errors are 0.43747 and 0.44031 meters respectively.

>>>>

\section{Notation}
We need to find out two functions, $f_x(x,y)$ and $f_y(x,y)$, to describe the
deviation. $x$ and $y$ are the true longitude and latitude respectively.
$f_x(x,y)$ and $f_y(x,y)$ are the deviation delta at point $(x,y)$.
In other words, if $(x,y)$ is the true coordinate, then
$(x+f_x(x,y),y+f_y(x,y))$ is the deviated coordinate.
Since $|f_x(x+d,y)-f_x(x,y)|$ is much smaller than $d$ (and same for $y$)
we can also
use the two functions to calculate the true coordinate from a deviated
coordinate.
Let $(x,y)$ be the deviated coordinate, $(x-f_x(x,y),y-f_y(x,y))$ is the
true coordinate.

We first solve the functions in one dimension by setting one parameter to be
constant.
I.e. instead of solving $f_x(x,y)$ directly, we first solve $f_x(x,c)$
and $f_x(c,y)$, where $c$ is a constant.
For example, we will look at $f_x(x,39.36)$ and $f_x(110.16,y)$.
Similarly, we will solve $f_y(x,c)$ and $f_y(c,y)$ before $f_y(x,y)$.

In the following analysis, we will use octave.
All the data used in the analysis has already been posted in my previous
article\footnote{url}.

@@

\noindent
\verb|> data=load("-ascii", "iny-f-110.txt");| \\
\verb|> y=data(:,2);| \\
\verb|> dx=data(:,3);| \\
\verb|> plot(y,dx)| \hfill Figure~\ref{fig:fxy-degree}(left)

@@

We can combine sin and cos of the same frequency.

\begin{verbatim}
> atan(0.82075/13.30099)
ans =  0.061628
> atan(0.27150/13.33287)
ans =  0.020360
> mx=[ones(size(x)),x,sin(2.*pi.*x-0.020360),sin(6.*pi.*x-0.061628)];
> std(mx*(pinv(mx)*dy)-dy)
ans =  0.38114
> pinv(mx)*dy
ans =
  -160.4519
     2.4276
    13.3356
    13.3263
\end{verbatim}
So the function is
\begin{eqnarray}
f_y(x,39.36) & = & -160.45195+2.42764x+13.33287sin(2\pi x) \nonumber \\
             &   & -0.27150cos(2\pi x)+13.30099cos(6\pi x)-0.82075cos(6\pi x) \nonumber \\
             & = & -160.45195+2.42764x+13.3356sin(2\pi x-0.061628) \nonumber \\
             &   & +13.3263cos(6\pi x-0.020360)
\end{eqnarray}

After trying out other constant latitudes, we find out that the first two
coefficients to be changing but the last 4 coefficients to be fixed.
Thus, we know
\begin{eqnarray}
f_y(x,y) & = & g_1(y)+g_2(y)x+13.3356sin(2\pi x-0.061628) \nonumber \\
         &   & +13.3263cos(6\pi x-0.020360)
\end{eqnarray}
where $g_1(y)$ and $g_2(y)$ are not yet known.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.6\textwidth]{fxx-lc2.png}
\end{center}
\caption{9-order polynomial (green) and 9-order polynomial + sinusoidal (blue)}
\label{fig:fxx-lc2}
\end{figure}

Although the 9-order polynomial fits the data well (error is 0.79 meters),
the function is ugly.
After lots of tries\footnote{
Most of the time has been spent here, but there is nothing to say about it
because it's simply trying different functions.
There are still a little bit of technique though.
For example, the two green peaks in Figure~\ref{fig:fxx-lc2} have distance
about 24 in x-axis, thus there might be a sinusoidal component of period 24.
}, I figured out that the low frequency
component actually consists of a quadratic curve and two sinusoidal wave of
period 60 and 24.
Using this model, the regression error is 0.39 meters and the function is
simple and beautiful.
\begin{verbatim}
> mx=[ones(size(x)),x,x.*x,sin(pi.*x./30),cos(pi.*x./30),\
  sin(pi.*x./12),cos(pi.*x./12),sin(pi.*x./3),cos(pi.*x./3),\
  sin(pi.*x),cos((pi.*x)),sin(2.*pi.*x),cos(2.*pi.*x),sin(6.*pi.*x),\
  cos((6.*pi.*x))]; std(mx*(pinv(mx)*dx)-dx)
ans =  0.38668
> pinv(mx)*dx
ans =
   1.2052e+03
  -1.8404e+01
   9.4022e-02
   8.3824e-01
   2.0044e+02
  -7.0435e+01
  -7.0449e+01
  -2.6699e+01
   9.1218e-02
  -1.3342e+01
   1.3044e-01
   1.3342e+01
  -2.7504e-01
   1.3306e+01
  -8.2298e-01
\end{verbatim}

@@

Finally, we have
\begin{eqnarray}
f_x(x,39.36) & = & 1205.1 -18.402x +0.094012x^2 +200.45sin(\pi x/30+1.5666) \nonumber \\
             &   & -99.62sin(\pi x/12+0.78550) -26.699sin(\pi x/3-0.0034165) \nonumber \\
             &   & -13.343sin(\pi x-0.0097763) +13.344sin(2\pi x-0.020612) \nonumber \\
             &   & +13.332sin(6\pi x-0.061772)
\end{eqnarray}
To see how the components fit the data, we can graduately add
the higher frequency components. Figure 5.5 shows this this idea. Figure 5.6 is
a zoomed in region in Figure 5.5. The first function (in red) is the quadratic
function $1205.1 -18.402x +0.094012x^2$. The second function (in green) adds
one sinusoidal component to it: $1205.1 -18.402x +0.094012x^2 +200.45sin(\pi
x/30+1.5666)$. The third function (in blue) adds one more sinusoidal component
to it, etc.
\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.48\textwidth]{fxx-lc3.png}
\includegraphics[width=0.48\textwidth]{fxx-lc4.png}
\end{center}
\caption{Decomposition of Components.
Left: overall; right: zoomed in.
The first function (red) has only the lowest frequency component.
The second function (green) adds the second lowest frequency component to the
first function.
The third (blue) adds another component, etc.}
\label{fig:fxx-lc3}
\end{figure}

After trying out other constant latitude, we found that only the first two
coefficients are affected by latitude. The rest of the coefficients do not
change. Thus, we have
\begin{eqnarray}
f_x(x,y) &=& g_1(y) -g_2(y)x +0.094012x^2 +200.45sin(\pi x/30+1.5666) \nonumber \\
         & & -99.62sin(\pi x/12+0.78550) -26.699sin(\pi x/3-0.0034165) \nonumber \\
         & & -13.343sin(\pi x-0.0097763) +13.344sin(2\pi x-0.020612) \nonumber \\
         & & +13.332sin(6\pi x-0.061772)
\end{eqnarray}
where $g_1(y)$ and $g_2(y)$ are not yet known.

@@

The process of finding $f_y(const, y)$ is similar to $f_x(x, const)$, except
that the two high frequency sinusoidal components do not appear in $f_y(const,
y)$.
\begin{verbatim}
> data=load("-ascii", "iny-a-110.txt");
> y=data(:,2);
> dy=data(:,4).*(6371000*pi/180);
> my=[ones(size(y)),y,y.*y,sin(pi.*y./30),cos(pi.*y./30),\
  sin(pi.*y./12),cos(pi.*y./12),sin(pi.*y./3),cos(pi.*y./3),\
  sin(pi.*y),cos((pi.*y))]; std(my*(pinv(my)*dy)-dy)
ans =    0.16683
> pinv(my)*dy
ans =
       162.1
     -22.422
     0.41569
     -172.78
      56.189
     -104.96
     -29.351
      13.251
      23.113
     -13.357
   -0.061311
> atan(56.189/-172.78)
ans =   -0.31442
> atan(-29.351/-104.96)
ans =    0.27267
> atan(23.113/13.251)
ans =     1.0502
> atan(-0.061311/-13.357)
ans =  0.0045901
> my=[ones(size(y)),y,y.*y,sin(pi.*y./30-0.31442),\
  sin(pi.*y./12+0.27267),sin(pi.*y./3+1.0502),\
  sin(pi.*y+0.0045901)]; std(my*(pinv(my)*dy)-dy)
ans =    0.16683
> pinv(my)*dy
ans =
      162.09
     -22.422
     0.41568
     -181.69
     -108.99
      26.643
     -13.357
\end{verbatim}

We have
\begin{eqnarray}
f_y(110.16,y) &=& 162.09-22.422y+0.41568y^2 \nonumber \\
              & & -181.69sin(\pi y/30-0.31442)-108.99sin(\pi y/12+0.27267) \nonumber \\
              & & +26.643sin(\pi y/3+1.05)-13.357sin(\pi y-0.00459) \\
f_y(x,y)      &=& g_1(x)+g_2(x)y+0.41568y^2 \nonumber \\
              & & -181.69sin(\pi y/30-0.31442)-108.99sin(\pi y/12+0.27267) \nonumber \\
              & & +26.643sin(\pi y/3+1.05)-13.357sin(\pi y-0.00459)
\end{eqnarray}

\section{The final 2D case: $f_x(x,y)$ and $f_y(x,y)$}
With out too much effort, we have found out the two functions:

\begin{eqnarray}
f_x(x,y) &=& 1598.1-23.605x -8.4345y +0.09933xy +0.10069x^2 \nonumber \\
          & & +199.1cos(\pi(x+0.017)/30) -99.772sin(\pi(x+0.0039)/12+\pi/4) \nonumber \\
          & & -26.687sin(\pi(x-0.0039)/3) -13.345sin(\pi(x-0.0039)) \nonumber \\
          & & +13.338sin(2\pi(x-0.0039)) +13.331sin(6\pi(x-0.0039)) \\
f_y(x,y) &=& -36.759 +1.5218x +10.002y -0.10032xy -0.03661y^2 \nonumber \\
         & & -217.44cos(\pi(y+8.143)/30)+106.41sin(\pi(y+1.1)/12) \nonumber \\
         & & -26.697sin(\pi(y+1)/3) +13.335sin(\pi y) \nonumber \\
         & & -13.329sin(2\pi(x-0.0039)) -13.319sin(6\pi(x-0.0039))
\end{eqnarray}
The standard errors are 0.43747 and 0.44031 meters respectively.
